---
title: "Titanic_Basic_Prediction"
output: word_document
---
# Titanic
## 1.Get Data
####  Download test and training set 

```{r}
setwd("C:/Users/sasla/Desktop/sezgi/Rprogramming/titanic")
train <- read.csv("train.csv", header = TRUE, stringsAsFactors = FALSE)
test <- read.csv("test.csv", header = TRUE, stringsAsFactors = FALSE)
```

####Import libraries
```{r}
library(ggplot2)
```

## 2.Basic Data exploration
#####PassengerId: and id given to each traveler on the boat
#####Pclass: the passenger class. It has three possible values: 1,2,3
#####The Name
#####The Sex
#####The Age
#####SibSp: number of siblings and spouses traveling with the passenger
#####Parch: number of parents and children traveling with the passenger
#####The ticket number
#####The ticket Fare
#####The cabin number
#####The embarkation. It has three possible values S,C,Q

```{r}
str(train)
names(train)
```

### a.Missing Data 

#### i-Use Check Missing Values for train
```{r pressure, echo=FALSE}
sapply(train, function(x) sum(is.na(x)))
sapply(train, function(x) sum(x==""))
```
##### As you see, there are missing values in Age,Cabin,Embarked

#### ii-Use Check Missing Values for test
```{r}
sapply(test, function(x) sum(is.na(x)))
sapply(test, function(x) sum(x==""))
#summary(test)
```
##### As you see, there are missing values in Age,Cabin,Fare

#### iii-Check NA values in Age to see if we remove or predict?
```{r}
sum(is.na(train$Age))/length(train$Age)
sum(train$Embarked=="")/length(train$Embarked)
sum(train$Cabin=="")/length(train$Cabin)

sum(test$Cabin=="")/length(test$Cabin)
sum(is.na(test$Age))/length(test$Age)
sum(is.na(test$Fare))/length(test$Fare)

```


#### See table of Survived
```{r}
table(train$Survived)
prop.table(table(train$Survived),1)
```

```{r}
summary(train$Sex)
for(i in c(3,5,7,12)){
  print( names(train)[i])
  print( prop.table(table(train[,i], train$Survived),1))
  barplot(table( train$Survived,train[,i]))
  boxplot(train$Survived,as.factor(train[,i]))
}
```

#### This shows %74 female survived and %81 man are dead 
#### According to sex and survived table we can set all females as survived"
#### Set all test$Survived values as 0 and let's try to submit to test.

```{r}
test$Survived <- 0
test$Survived[test$Sex == 'female'] <- 1
submit <- data.frame(PassengerId = test$PassengerId, Survived = test$Survived)
write.csv(submit, file = "first.csv", row.names = FALSE)